
# Importing and Loading Packages

```{r, results = FALSE, message=FALSE, warning=FALSE}


install.packages("tidyverse", repos = "http://cran.us.r-project.org")
install.packages("ggwordcloud", repos = "http://cran.us.r-project.org")
install.packages("viridis", repos = "http://cran.us.r-project.org")
install.packages("scales", repos = "http://cran.us.r-project.org")

library(tidyverse)
library(ggwordcloud)
library(viridis)
library(scales)
```


# Loading Datasets

```{r}
#Main Data set
ds_df <- read.csv("ds_salaries.csv")

#Country codes Data set, used for cleaning later
iso <- read.csv("wikipedia-iso-country-codes.csv")
```


# Inspecting the Data

```{r}
#Inspecting the dataset

colnames(ds_df)  #List of column names
```

```{r}
nrow(ds_df)  #How many rows are in the data frame?
```

```{r}
dim(ds_df)   #Dimensions of the data frame
```

```{r}
head(ds_df)  #Displays the first couple of rows of the data frame
```

```{r}
str(ds_df)  #Display list of columns and data types
```

```{r}
summary(ds_df)  #Statistical summary of data 
```

```{r}
# Checking for any NA values
any(is.na(ds_df))
```
```{r}

#Checking for unique values per column
ds_df %>%
  summarise(work_year = n_distinct(work_year), experience_level = n_distinct(experience_level), employment_type = n_distinct(employment_type), job_title = n_distinct(job_title), salary = n_distinct(salary), salary_currency = n_distinct(salary_currency), salary_in_usd = n_distinct(salary_in_usd), employee_residence = n_distinct(employee_residence), remote_ratio = n_distinct(remote_ratio), company_location = n_distinct(company_location), company_size = n_distinct(company_size) )

```

# Cleaning the Data

#### There are a few problems we need to fix:

##### 1. The columns experience_level, employment_type, employee_residence, company_size and company_location contain abbreviations that may not be understood to some, we will rename the data contained in those columns to more descriptive labels.

```{r}
#Renamed abbreviations for the experience_level column

ds_df$experience_level[ds_df$experience_level == "SE"] <- "Senior-level"
ds_df$experience_level[ds_df$experience_level == "MI"] <- "Indermediate-level"
ds_df$experience_level[ds_df$experience_level == "EN"] <- "Junior/Entry-level"
ds_df$experience_level[ds_df$experience_level == "EX"] <- "Director/Executive-level"

#Renamed abbreviations for the employment_type column

ds_df$employment_type[ds_df$employment_type == "FT"] <- "Full Time"
ds_df$employment_type[ds_df$employment_type == "PT"] <- "Part Time"
ds_df$employment_type[ds_df$employment_type == "CT"] <- "Contract"
ds_df$employment_type[ds_df$employment_type == "FL"] <- "Free Lance"

#Renamed abbreviations for the company_size column

ds_df$company_size[ds_df$company_size == "S"] <- "Small"
ds_df$company_size[ds_df$company_size == "M"] <- "Medium"
ds_df$company_size[ds_df$company_size == "L"] <- "Large"


#Pulling columns from ds_df and iso into vectors that we can iterate from.

employee_residence <- ds_df %>% pull(employee_residence)

Alpha.2.code <- iso %>% pull(Alpha.2.code)

Country_name <- iso %>% pull(English.short.name.lower.case)

index <- 0
new_employee_residence <- c()

#Looping through each item in new_employee_residence

for (item in employee_residence) {
  
#check where the item exists in Alpha.2.code, then assign the index found in Alpha.2.code to index

    index <- which(Alpha.2.code == item)[1]
    
#use the index to find the corresponding Country_name, then append that Country_name to new_employee_residence, save to new_employee_residence
    
    new_employee_residence <- append(new_employee_residence, Country_name[index])
}

# assign new_employee_residence to the employee_residence column

ds_df$employee_residence <- new_employee_residence

```

```{r}
str(ds_df$experience_level)
str(ds_df$employment_type)
str(ds_df$employee_residence)
str(ds_df$company_size)
```

```{r}

#Pulling company_location from ds_df into a vector that we can iterate from.

company_location <- ds_df %>% pull(company_location)

index <- 0

new_company_location <- c()


#Same For loop as above, but for company_location
for (item in company_location) {
    index <- which(Alpha.2.code == item)[1]
    new_company_location <- append(new_company_location, Country_name[index])
}

ds_df$company_location <- new_company_location

ds_df$employee_residence[ds_df$employee_residence == "United States Of America"] <- "USA"
ds_df$employee_residence[ds_df$employee_residence == "United Kingdom"] <- "UK"

ds_df$company_location[ds_df$company_location == "United States Of America"] <- "USA"
ds_df$company_location[ds_df$company_location == "United Kingdom"] <- "UK"

```

```{r}
str(ds_df$company_location)
```

##### 2. We do not need the salary or salary_currency columns for our analysis.


```{r}
ds_df <- ds_df %>%
  select(-c(salary, salary_currency))

str(ds_df)
```

##### 3. Since remote_ratio contains ratios they all should have % at the end.

```{r}

#Add a % at the end of every element in remote_ratio
remote_ratio <- ds_df %>% pull(remote_ratio)
new_remote_ratio <- c()
for(item in remote_ratio) {
    new_remote_ratio <- append(new_remote_ratio, paste(item, "%", sep = ""))
}

ds_df$remote_ratio <- new_remote_ratio

```

```{r}
str(ds_df$remote_ratio)
```

# Analysis

```{r}
# What's the distribution of experience_level?
experience_level <- data.frame(experience_level = ds_df$experience_level)

ggplot(data = experience_level) +
     geom_bar(mapping = aes(x = experience_level, fill = experience_level))+
     theme(axis.text.x = element_text(angle =20, hjust = 1))+
     labs(title = "Distribution of Experience Level")
```
  
Most people with data science related jobs are Senior level.  

```{r}
query <- ds_df %>%
  select(job_title, experience_level) %>%
  group_by(job_title, experience_level) %>%
  summarise("Count" = n()) %>%
  arrange(desc(Count)) %>%
  filter(job_title == "Data Engineer" |job_title == "Data Scientist" |job_title == "Data Analyst" |job_title == "Machine Learning Engineer" |job_title == "Analytics Engineer" |job_title == "Data Architect" |job_title == "Research Scientist" |job_title == "Applied Scientist" |job_title == "Data Science Manager" |job_title == "Research Engineer")



ggplot(query, aes(x = job_title, y = Count, fill = experience_level)) +
  geom_col()+
  theme(axis.text.x = element_text(angle =40, hjust = 1))+
  labs(title = "Experience Level per Job Title", subtitle = "For the 10 most popular data science jobs")

```
The ratio of Seniors vs other experience levels is roughly the same per job title.

```{r}
employment_type <- data.frame(employment_type = ds_df$employment_type)

ggplot(data = employment_type) +
    geom_bar(mapping = aes(x = employment_type, fill = employment_type))+
    labs(title = "Employment type distribution")
```
  
Virtually all employees are full time.  
```{r}
job_title <- data.frame(table(job_title = ds_df$job_title))

ggplot(data = job_title, aes(label = job_title, size = Freq, color = job_title)) +
     geom_text_wordcloud(rm_outside = TRUE) +
     scale_size_area(max_size = 20) +
     theme_minimal()+
     labs(title = "Word Cloud of Data Science Jobs")
```

```{r}
salary <- data.frame(table(ds_df$salary_in_usd))

salary$Var1 <- as.numeric(as.character(salary$Var1))

ggplot(salary, aes(x = Var1)) +
     geom_histogram(binwidth = 50000, fill="lightblue", color="black")+
     scale_x_continuous(labels = scales::comma)+
     labs(title = "Distribution of Salary", x = "Salary", y = "Count")

print(summary(salary))
```
  
The average salary for all Data Science related jobs is 130000\$, with half of all salaries lie between 68000\$ and 180000\$  
```{r}



 data <- ds_df %>%
     select(job_title, salary_in_usd,
     employee_residence) %>%
     filter(job_title == "Data Analyst" | job_title == "BI Analyst" | job_title == "BI Data Analyst" | job_title == "Business Data Analyst" | job_title == "Data Analytics Specialist" | job_title == "Data Operations Analyst" | job_title == "Data Quality Analyst" | job_title == "Finance Data Analyst" | job_title == "Financial Data Analyst" | job_title == "Insight Analyst" | job_title == "Lead Data Analyst" | job_title == "Marketing Data Analyst" | job_title == "Principal Data Analyst" | job_title == "Product Data Analyst" | job_title == "Staff Data Analyst")%>%
     group_by(employee_residence) %>%
     summarise(mean = mean(salary_in_usd))
     
     
world <- map_data("world")

world %>%
     merge(data, by.x = "region", by.y = "employee_residence", all.x = T) %>%
     arrange(group, order) %>%
     ggplot(aes(x = long, y = lat, group = group, fill = mean))+ 
     geom_polygon(color = "white", linewidth = 0.2) +
     scale_fill_viridis("", na.value = "gray90", option = "C") +
     theme_minimal() +
     theme(axis.text = element_blank(),
           axis.title = element_blank(),
           
           
           panel.grid = element_blank())+
     labs(title = "Average Salary of Data Analysts per Country")
```
  
Data Analysis in China make the most, followed by data analysts in the US.  
```{r}
ds_df$salary_in_usd <- as.numeric(ds_df$salary_in_usd)

temp <- data.frame(table(ds_df$job_title))

temp <- temp %>%
  arrange(desc(Freq)) %>%
  head(n = 70)

data <- ds_df %>%
  select(job_title, salary_in_usd) %>%
  filter(job_title == "Data Engineer" | job_title == "Data Scientist" | job_title == "Data Analyst" | job_title == "Machine Learning Engineer" | job_title == "Analytics Engineer" | job_title == "Data Architect" | job_title == "Research Scientist" | job_title == "Applied Scientist" | job_title == "Data Science Manager" | job_title == "Research Engineer" | job_title == "ML Engineer" | job_title == "Data Manager" | job_title == "Machine Learning Scientist" | job_title == "Data Science Consultant" | job_title == "Data Analytics Manager" | job_title == "Computer Vision Engineer" | job_title == "AI Scientist" | job_title == "BI Data Analyst" | job_title == "Business Data Analyst" | job_title == "Data Specialist" | job_title == "BI Developer" | job_title == "Applied Machine Learning Scientist" | job_title == "AI Developer" | job_title == "Big Data Engineer" | job_title == "Director of Data Science" | job_title == "Machine Learning Infrastructure Engineer" | job_title == "Applied Data Scientist" | job_title == "Data Operations Engineer" | job_title == "ETL Developer" | job_title == "Head of Data")

data %>%
     ggplot(aes(job_title, salary_in_usd, fill = job_title))+
     geom_boxplot()+
     theme_bw()+
     theme(axis.text.x = element_text(angle = 90), legend.position = "none")+
     scale_y_continuous(labels = comma)+
     labs(title = "Salary Ranges for Data Science Jobs", subtitle = "30 most popular data science jobs")

```

```{r}
data <- data.frame(table(ds_df$employee_residence))

world <- map_data("world")

world %>%
     merge(data, by.x = "region", by.y = "Var1", all.x = T) %>%
     arrange(group, order) %>%
     ggplot(aes(x = long, y = lat, group = group, fill = Freq))+
     geom_polygon(color = "white", linewidth = 0.2) +
     scale_fill_viridis("", na.value = "gray90", option = "C") +
     theme_minimal() +
     theme(axis.text = element_blank(),
           axis.title = element_blank(),
           panel.grid = element_blank())+
     labs(title = "Employee Residence")
```
  
Virtually all employees in the data live in the US.  

```{r}
remote_ratio <- data.frame(table(ds_df$remote_ratio))

remote_ratio %>%
     ggplot(aes(x = Var1, y = Freq))+
     geom_col(fill = "lightblue", color = "black")+
     xlab("Remote Ratio") +
     ylab("Count")+
     labs(title = "Distribution of Remote Ratio")
```

```{r}
data <- data.frame(table(ds_df$company_location))

world <- map_data("world")

world %>%
     merge(data, by.x = "region", by.y = "Var1", all.x = T) %>%
     arrange(group, order) %>%
     ggplot(aes(x = long, y = lat, group = group, fill = Freq))+
     geom_polygon(color = "white", linewidth = 0.2) +
     scale_fill_viridis("", na.value = "gray90", option = "C") +
     theme_minimal() +
     theme(axis.text = element_blank(),
           axis.title = element_blank(),
           panel.grid = element_blank())+
     ggtitle("Company Location")
```
  
Virtually all data science companies are located in the US.  



```{r}
company_size <- data.frame(table(ds_df$company_size))

 company_size %>%
     ggplot(aes(x = Var1, y = Freq))+
     geom_col(fill = "lightblue", color = "black")+
     labs(title = "Company Size Distribution", x = "Company Size", y = "Count")
```
  
Most Companies are medium sized (50-250 employees).  

# Conclusions

In this analysis, we explored the salaries of data science jobs in the US using data from Kaggle. We found that the average salary of all data related jobs is 130,000\$, and half of all salaries lie between 68000\$ and 180000\$.We also identified some trends and patterns in the data science job market, such as most employee's experience being senior-level and most data science companies are located in the US with company sizes ranging from 50-250 employees.


In conclusion, our analysis provides a valuable snapshot of the current state and trends of data science salaries in the US, which can help data scientists and employers make informed decisions and plan their careers.


